---
title: "Project 1"
author: "Zichang Xiang"
date: "6/16/2021"
output:  
  github_document:  
    toc: true
---
### Reading and summarizing data from NHL API
```{r setup, include=FALSE}
#install.packages("rmarkdown")
#load the required packages
library(rmarkdown)
library(dplyr)
library(jsonlite)
library(RCurl)
library(httr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

#### Required Packages
jsonlite, RCurl, rmarkdown, httr, and ggplot2

```{r}
#load the required packages
library(rmarkdown)
library(dplyr)
library(jsonlite)
library(RCurl)
library(httr)
library(ggplot2)
```

### functions to contact the NHL records API

#### function to return franchise
```{r franchise}
#retrieve the info and convert to a list
get_franchise <- GET("https://records.nhl.com/site/api/franchise")
franchise_cont <- content(get_franchise, "text", encoding = "UTF-8")
franchise_json<- fromJSON(franchise_cont, flatten = TRUE)
franchise_list <-as_tibble(franchise_json$data)
#create function
franchise <- function(name) {
  if (is.null(name)){
    return(franchise_list)
  }else if (is.character(name)){
  return(franchise_list %>% filter(fullName == name | teamAbbrev == name | teamCommonName == name |teamPlaceName ==    name) %>% collect())
  }else
  return(franchise_list %>% filter(mostRecentTeamId == name) %>% collect())
}
franchise(NULL)
```


#### function to return total stats for every franchise
```{r total}
#retrieve info and convert to a list
get_total <- GET("https://records.nhl.com/site/api/franchise-team-totals")
total_cont <- content(get_total, "text", encoding = "UTF-8")
total_json<- fromJSON(total_cont, flatten = TRUE)
total_list <-as_tibble(total_json$data)
#create function team_totals
team_totals <- function(name) {
  if (is.null(name)){
    return(total_list)
  }else if (is.character(name)){
  return(total_list %>% filter(teamName == name | triCode == name) )
  }else
  return(total_list %>% filter(franchiseId == name) )
}

team_totals("MTL")
team_totals(NULL)
```

#### function to return season records
```{r season}
#create function season
season <- function(name) {
  if (is.null(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-season-records")
  get_season <- GET(base_url)
  season_cont <- content(get_season, "text", encoding = "UTF-8")
  season_json <- fromJSON(season_cont, flatten = TRUE)
  season_list <- as_tibble(season_json$data)
  return(season_list)
  }else if (is.numeric(name)){
   base_url <- paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", name)
   get_season <- GET(base_url)
   season_cont <- content(get_season, "text", encoding = "UTF-8")
   season_json <- fromJSON(season_cont, flatten = TRUE)
   season_list <- as_tibble(season_json$data)
  return(season_list)}else{
  get_season <- GET("https://records.nhl.com/site/api/franchise-season-records")
  season_cont <- content(get_season, "text", encoding = "UTF-8")
  season_json <- fromJSON(season_cont, flatten = TRUE)
  season_list <- as_tibble(season_json$data)
  return(season_list %>% filter(franchiseName == name))
}}
season("Montréal Canadiens")
season(NULL)
```
#### function to return goalie records
```{r goalie}
#create function goalie
goalie <- function(name) {
  if (is.null(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records")
  get_goalie <- GET(base_url)
  goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
  goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
  goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list)
  }else if (is.numeric(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", name)
  get_goalie <- GET(base_url)
  goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
  goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
  goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list %>% filter(franchiseId==name))}else{
  get_goalie <- GET("https://records.nhl.com/site/api/franchise-goalie-records")
  goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
  goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
  goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list %>% filter(franchiseName == name|lastName==name|firstName==name))
}}
goalie(1)
goalie(NULL)
```
#### function to return skater records
```{r}
#create function skater
skater <- function(name) {
  if (is.null(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records")
  get_skater <- GET(base_url)
  skater_cont <- content(get_skater, "text", encoding = "UTF-8")
  skater_json <- fromJSON(skater_cont, flatten = TRUE)
  skater_list <- as_tibble(skater_json$data)
  return(skater_list)
  }else if (is.numeric(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", name)
  get_skater <- GET(base_url)
  skater_cont <- content(get_skater, "text", encoding = "UTF-8")
  skater_json <- fromJSON(skater_cont, flatten = TRUE)
  skater_list <- as_tibble(skater_json$data)
  return(skater_list %>% filter(franchiseId==name))}else{
  get_skater<- GET("https://records.nhl.com/site/api/franchise-skater-records")
  skater_cont <- content(get_skater, "text", encoding = "UTF-8")
  skater_json <- fromJSON(skater_cont, flatten = TRUE)
  skater_list <- as_tibble(skater_json$data)
  return(skater_list %>% filter(franchiseName == name|lastName==name|firstName==name))
}}
skater(1)
skater(NULL)
```

#### function to return detail records
```{r detail}
#create function detail
detail <- function(name) {
  if (is.null(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-detail?")
  get_detail <- GET(base_url)
  detail_cont <- content(get_detail, "text", encoding = "UTF-8")
  detail_json <- fromJSON(detail_cont, flatten = TRUE)
  detail_list <- as_tibble(detail_json$data)
  return(detail_list)  
  }else if (is.numeric(name)){
  base_url <- paste0("https://records.nhl.com/site/api/franchise-detail?cayenneExp=mostRecentTeamId=", name)
  get_detail <- GET(base_url)
  detail_cont <- content(get_detail, "text", encoding = "UTF-8")
  detail_json <- fromJSON(detail_cont, flatten = TRUE)
  detail_list <- as_tibble(detail_json$data)
  return(detail_list %>% filter(mostRecentTeamId==name))}else{
  get_detail <- GET("https://records.nhl.com/site/api/franchise-detail")
  detail_cont <- content(get_detail, "text", encoding = "UTF-8")
  detail_json <- fromJSON(detail_cont, flatten = TRUE)
  detail_list <- as_tibble(detail_json$data)
  return(detail_list %>% filter(teamAbbrev == name|teamFullName==name))
}}
detail("Vancouver Canucks")
detail(NULL)
```

### function to contact the NHL stats API
```{r stats}
#create function stats
stats <- function(name) {
  if (is.null(name)){
    full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams","?expand=team.stats")
    }else{full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", name, "/?expand=team.stats")}
    get_stats <- GET(full_url)
    stats_cont <- content(get_stats, "text", encoding = "UTF-8")
    stats_json <- fromJSON(stats_cont, flatten = TRUE)
    stats_list <- as_tibble(stats_json$teams)
    return(stats_list)
}
stats(NULL)
```

### Wrapper function to call other function
```{r}
#create wrapper function to call other functions
wrapper <- function(fun, name){
  if (fun == "franchise"){
    return(franchise(name))
  }else if (fun == "team_totals"){
    return(team_totals(name))
  }else if (fun == "season"){
    return(season(name))
  }else if (fun == "goalie"){
    return(goalie(name))
  }else if (fun == "skater"){
    return(skater(name))
  }else if (fun == "detail"){
    return(detail(name))
  }else if (fun == "stats"){
    return(stats(name))
  }else{
    stop("Type the correct function!")
  }
  }
wrapper("stats", NULL)
wrapper("franchise", NULL)
```

### Basic exploratory data analysis
#### Combine data from two endpoints
```{r}
#all goalies
goalie_all <- goalie(NULL)[,c(1:8)]
goalie_all <- goalie_all %>% mutate(Type = "goalie")
goalie_all
#all skaters
skater_all <-skater(NULL)[,c(1:2,4:8,10)]
skater_all <- skater_all %>% mutate(Type = "skater")
skater_all
#all goalies and skaters 
goalie_skater_all <- rbind(goalie_all, skater_all)
goalie_skater_all <- goalie_skater_all[order(goalie_skater_all$id),]
goalie_skater_all
```

#### create at least two new variables
```{r}
#all goalies
goalie_all <- goalie(NULL)[,c(1:9, 28:29)]
goalie_all
#create three new variables
goalie_all <- goalie_all %>% mutate(loss_rate = goalie_all$losses/goalie_all$gamesPlayed, win_rate = goalie_all$wins/goalie_all$gamesPlayed, tie_rate =goalie_all$ties/goalie_all$gamesPlayed)
goalie_all
```


#### create contingency tables
```{r}
#table to summarize the relationship between player type and the number of active player
table(goalie_skater_all$Type, goalie_skater_all$activePlayer)
#table to summarize the relationship between divisions and timeZones
table(stats(NULL)$division.name, stats(NULL)$venue.timeZone.tz)
```


#### create numerical summaries
```{r}
#quartiles and mean for wins in regular season
totals_regular <- team_totals(NULL) %>% filter(gameTypeId ==2) 
summary(totals_regular$wins)
#quartiles and mean for games played
goalie_skater_all %>% group_by(Type) %>% summarize(min = min(gamesPlayed), avg = round(mean(gamesPlayed),0), med = median(gamesPlayed), max = max(gamesPlayed))
```

#### creates plots
```{r}
#create bar plot to compare number of players for each player type
g <- ggplot(goalie_skater_all, aes(x=Type))
g + geom_bar(aes(fill=as.factor(activePlayer)), position = "dodge")+ 
  ggtitle("Numer of Players for Each Player Type")+ 
  labs(x="Player Type")+
  scale_fill_discrete(name = "Active player")
```


```{r}
#create histogram
g <- ggplot(goalie_all, aes(x=wins, ..density..))
g + geom_histogram(bins  = 150)+
  labs(x="Wins", y="Density")+
  ggtitle("Histogram for Wins")+
  geom_density(col="red", lwd = 3, adjust =0.4)
```


```{r}
#create two new datasets
conference <- stats(NULL)[,c("id", "conference.name")]
totals <- team_totals(NULL) %>% filter(gameTypeId == 2, teamId %in% c(1:26,28:30,52:55)) %>% select(c(14, 26,27,30))
totals <- totals %>% rename(id = teamId)
conference <- conference %>% rename(Conference = conference.name)
conf_totals<- merge(totals,conference, by = "id")
#create boxplot for number of wins for each conference
g <- ggplot(conf_totals, aes(x = Conference, y = wins)) 
g + geom_boxplot() + 
  geom_jitter(aes(color = Conference)) + 
  ggtitle("Boxplot for Wins")+
  labs(x="Conference", y= "Wins")
conf_totals
```

```{r}
#create scatter plot
g <- ggplot(conf_totals, aes(x = wins, y = losses, group = Conference)) 
g+ geom_point(aes(color=Conference), position="jitter")+
   geom_smooth(aes(group=Conference),method=lm, formula= y~x, col="green")+
   ggtitle("Wins vs Losses")
```

