---
title: "Project 1"
author: "Zichang Xiang"
date: "6/16/2021"
purpose: create vignette for reading and summarizing data from NHL API
output:  
  github_document:  
    toc: true
---
### Reading and summarizing data from NHL API
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Required Packages
rmarkdown, jsonlite, ggplot2, dplyr, knitr, tidyr, RCurl, httr

```{r message = FALSE}
#load the required packages
library(rmarkdown)
library(jsonlite)
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(RCurl)
library(httr)
```

### Functions to contact the NHL records API
#### Function to return franchise
```{r franchise}
#retrieve the info and convert to a list
get_franchise  <- GET("https://records.nhl.com/site/api/franchise")
franchise_cont <- content(get_franchise, "text", encoding = "UTF-8")
franchise_json <- fromJSON(franchise_cont, flatten = TRUE)
franchise_list <- as_tibble(franchise_json$data)
franchise_list
#create data frame lookup_f
lookup_f <- franchise_list[ , c(3,5)]
lookup_f

#create function franchise
franchise <- function(name) {
  if (is.null(name)){
    return(franchise_list)
  }else if (name %in% lookup_f$mostRecentTeamId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise?cayenneExp=mostRecentTeamId=", name)
  }else if (toupper(name) %in% toupper(lookup_f$fullName)){
    id <- lookup_f %>% filter(toupper(fullName) == toupper(name)) %>% select(2)
    full_url <- paste0("https://records.nhl.com/site/api/franchise?cayenneExp=mostRecentTeamId=", id)
  }else{
    stop("Wrong input!")
  }
    get_franchise  <- GET(full_url)
    franchise_cont <- content(get_franchise, "text", encoding = "UTF-8")
    franchise_json <- fromJSON(franchise_cont, flatten = TRUE)
    franchise_list <- as_tibble(franchise_json$data)
    return(franchise_list)
}

#try to use function franchise
franchise(8)
franchise("MOntréal Canadiens")

```


#### Function to return total stats for every franchise
```{r total}
#retrieve info and convert to a list
get_total  <- GET("https://records.nhl.com/site/api/franchise-team-totals")
total_cont <- content(get_total, "text", encoding = "UTF-8")
total_json <- fromJSON(total_cont, flatten = TRUE)
total_list <- as_tibble(total_json$data)
total_list
#create data frame lookup_t
lookup_t <- total_list[ , c(4,27)]
lookup_t

#create function total
total <- function(name) {
  if (is.null(name)){
    return(total_list)
  }else if (name %in% lookup_t$franchiseId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise-team-totals?cayenneExp=franchiseId=", name)
  }else if (toupper(name) %in% toupper(lookup_t$teamName)){
    id <- lookup_t %>% filter(toupper(teamName) == toupper(name)) %>% select(1) %>% unique()
    full_url <- paste0("https://records.nhl.com/site/api/franchise-team-totals?cayenneExp=franchiseId=", id)
  }else{
    stop("Wrong input!")
  }
    get_total  <- GET(full_url)
    total_cont <- content(get_total, "text", encoding = "UTF-8")
    total_json <- fromJSON(total_cont, flatten = TRUE)
    total_list <- as_tibble(total_json$data)
    return(total_list)
}

#try to use function total
total(1)
total("NEw York Rangers")

```

#### Function to return season records
```{r season}
#create data frame lookup_s including name and id
base_url    <- paste0("https://records.nhl.com/site/api/franchise-season-records")
get_season  <- GET(base_url)
season_cont <- content(get_season, "text", encoding = "UTF-8")
season_json <- fromJSON(season_cont, flatten = TRUE)
season_list <- as_tibble(season_json$data)
lookup_s <- season_list[,c(14,15)]
lookup_s

#create function season
season <- function(name) {
  if (is.null(name)){
    return(season_list)
  }else if (name %in% lookup_s$franchiseId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", name)
  }else if (toupper(name) %in% toupper(lookup_s$franchiseName)){
    id <- lookup_s %>% filter(toupper(franchiseName) == toupper(name)) %>% select(1)
    full_url <- paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", id)
  }else{
    stop("Wrong input!")
  }
    get_season  <- GET(full_url)
    season_cont <- content(get_season, "text", encoding = "UTF-8")
    season_json <- fromJSON(season_cont, flatten = TRUE)
    season_list <- as_tibble(season_json$data)
    return(season_list)
}

#try to use function season
season(1)
season("MOntréal Canadiens")
```

#### Function to return goalie records
```{r goalie}
#create data frame lookup_g
base_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records")
get_goalie  <- GET(base_url)
goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
goalie_list <- as_tibble(goalie_json$data)
lookup_g <- goalie_list[ , c(4, 5)]
lookup_g

#create function goalie
goalie <- function(name) {
  if (is.null(name)){
    return(goalie_list)
  }else if (name %in% lookup_g$franchiseId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", name)
  }else if (toupper(name) %in% toupper(lookup_g$franchiseName)){
    id <- lookup_g %>% filter(toupper(franchiseName) == toupper(name)) %>% select(1) %>% unique()
    full_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", id)
  }else{
    stop("Wrong input!")
  }
    get_goalie  <- GET(full_url)
    goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
    goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
    goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list)
}

#try to use function goalie
goalie(15)
goalie("DAllas Stars")
```

#### Function to return skater records
```{r}
#create data frame lookup_sk
base_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records")
get_skater  <- GET(base_url)
skater_cont <- content(get_skater, "text", encoding = "UTF-8")
skater_json <- fromJSON(skater_cont, flatten = TRUE)
skater_list <- as_tibble(skater_json$data)
lookup_sk <- skater_list[ , c(5, 6)]
lookup_sk

#create function skater
skater <- function(name) {
  if (is.null(name)){
    return(skater_list)
  }else if (name %in% lookup_sk$franchiseId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", name)
  }else if (toupper(name) %in% toupper(lookup_sk$franchiseName)){
    id <- lookup_sk %>% filter(toupper(franchiseName) == toupper(name)) %>% select(1) %>% unique()
    full_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", id)
  }else{
    stop("Wrong input!")
  }
    get_skater  <- GET(full_url)
    skater_cont <- content(get_skater, "text", encoding = "UTF-8")
    skater_json <- fromJSON(skater_cont, flatten = TRUE)
    skater_list <- as_tibble(skater_json$data)
  return(skater_list)
}

#try to use function skater
skater(1)
skater("WAshington Capitals")
```

#### Function to return detail records
```{r detail}
#create data frame lookup_d
base_url <- paste0("https://records.nhl.com/site/api/franchise-detail?")
get_detail  <- GET(base_url)
detail_cont <- content(get_detail, "text", encoding = "UTF-8")
detail_json <- fromJSON(detail_cont, flatten = TRUE)
detail_list <- as_tibble(detail_json$data)
lookup_d <- detail_list[, c(10,13)]
lookup_d

#create function detail
detail<- function(name) {
  if (is.null(name)){
    return(detail_list)
  }
  if (name %in% lookup_d$mostRecentTeamId){
    full_url <- paste0("https://records.nhl.com/site/api/franchise-detail?cayenneExp=mostRecentTeamId=", name)
  }else if (toupper(name) %in% toupper(lookup_d$teamFullName)){
    id <- lookup_d %>% filter(toupper(teamFullName) == toupper(name)) %>% select(1) %>% unique()
    full_url <- paste0("https://records.nhl.com/site/api/franchise-detail?cayenneExp=mostRecentTeamId=", id)
  }else{
    stop("Wrong input!")
  }
    get_detail  <- GET(full_url)
    detail_cont <- content(get_detail, "text", encoding = "UTF-8")
    detail_json <- fromJSON(detail_cont, flatten = TRUE)
    detail_list <- as_tibble(detail_json$data)
  return(detail_list)
}

#try to use function detail
detail(15)
detail("DAllas Stars")
```


### Function to contact the NHL stats API
```{r stats}
#create data frame lookup_st
full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams", "?expand=team.stats")
get_stats  <- GET(full_url)
stats_cont <- content(get_stats, "text", encoding = "UTF-8")
stats_json <- fromJSON(stats_cont, flatten = TRUE)
stats_list <- as_tibble(stats_json$teams)
lookup_st <- stats_list[, c(1:2)]
lookup_st <- rename(lookup_st, teamName = name)
lookup_st

#create function stats
stats <- function(name) {
  if (is.null(name)){
    return(stats_list)
  }else if (name %in% lookup_st$id){
    full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", name,
                           "/?expand=team.stats")
  }else if (toupper(name) %in% toupper(lookup_st$teamName)){
    id <- lookup_st %>% filter(toupper(teamName) == toupper(name)) %>% select(1)
    full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", id,
                           "/?expand=team.stats")
  }else{
    stop("Wrong input!")
  }
   get_stats  <- GET(full_url)
   stats_cont <- content(get_stats, "text", encoding = "UTF-8")
   stats_json <- fromJSON(stats_cont, flatten = TRUE)
   stats_list <- as_tibble(stats_json$teams)
#   teamStats <- stats_list$teamStats %>% unnest(teamStats) %>% unnest(splits)
  return(stats_list %>% unnest(c(teamStats)) %>% unnest(c(splits)))
}

#try to use function stats
stats(15)
stats("DAllas Stars")
```

### Wrapper function to call other functions
```{r}
#create wrapper function to call other functions
wrapper <- function(fun, name){
  if (fun == "franchise"){
    return(franchise(name))
  }else if (fun == "total"){
    return(total(name))
  }else if (fun == "season"){
    return(season(name))
  }else if (fun == "goalie"){
    return(goalie(name))
  }else if (fun == "skater"){
    return(skater(name))
  }else if (fun == "detail"){
    return(detail(name))
  }else if (fun == "stats"){
    return(stats(name))
  }else{
    stop("Type the correct function!")
  }
}

#try to use function wrapper
wrapper("stats", 1)
wrapper("franchise", 8)
```

### Basic exploratory data analysis
#### Combine data from two endpoints
```{r}
#create new data frame goalie_all
goalie_all <- goalie(NULL)[ , c(1:8)]
goalie_all <- goalie_all %>% mutate(Type = "goalie")
#view the data frame
goalie_all

#create new data frame skater_all
skater_all <- skater(NULL)[ , c(1:2, 4:8, 10)]
skater_all <- skater_all %>% mutate(Type = "skater")
#view the data frame
skater_all

#create new data frame goalie_skater_all
goalie_skater_all <- rbind(goalie_all, skater_all)
goalie_skater_all <- goalie_skater_all[order(goalie_skater_all$id), ]
#view the data frame
goalie_skater_all
```

#### Create at least two new variables  
Here we create three new variables loss_rate, win_rate, and tie_rate
```{r}
#create new data frame goalie_all
goalie_all <- goalie(NULL)[ , c(1:9, 28:29)]
#view the data frame
goalie_all

#create three new variables loss_rate, win_rate, and tie_rate
goalie_all <- goalie_all %>% mutate(loss_rate = goalie_all$losses/goalie_all$gamesPlayed, 
                                    win_rate = goalie_all$wins/goalie_all$gamesPlayed, 
                                    tie_rate = goalie_all$ties/goalie_all$gamesPlayed)
#view the data frame
goalie_all
```

#### Create contingency tables  
Table 1 below shows the relationship between the type of player and the number of active players. As we can see from the table, there are more non-active players than active players. This is true for both goalies and skaters.

In the second table, we can find the number of teams for each combination of division and time zone. We can see from the table that teams in division Honda West and in division Scotia North are from four different time zones, while teams in division MassMutual East fall into only one time zone EDT.
```{r}
#create table to summarize the relationship between player type and the number of active players
table_1 <- table(goalie_skater_all$Type, goalie_skater_all$activePlayer)
kable(table_1)
#create table to summarize the relationship between divisions and timeZones
table_2 <-table(stats_list$division.name, stats_list$venue.timeZone.tz)
kable(table_2)
```

#### Create numerical summaries  
Table 1 shows the numerical summaries of wins in regular season. We notice that the median number of wins is 678 in regular season.
Table 2 show the numerical summaries for different type of players. We notice that the average number of games played by skater is higher than that by goalie.
```{r}
#quantiles and mean for wins in regular season
totals_regular <- total(NULL) %>% filter(gameTypeId ==2) 
summary(totals_regular$wins)

#quantiles and mean for games played
goalie_skater_all %>% group_by(Type) %>% summarize(min = min(gamesPlayed), 
                                                   avg = round(mean(gamesPlayed),0), 
                                                   med = median(gamesPlayed), 
                                                   max = max(gamesPlayed))
```

#### Create plots  
The bar plot below compares the number of players for each player type. As we can see, non-active skaters are the most common, while active goalies are the least common.
```{r}
#create bar plot to compare number of players for each player type
g <- ggplot(goalie_skater_all, aes(x = Type))
g + geom_bar(aes(fill = as.factor(activePlayer)), position = "dodge") + 
  ggtitle("Numer of Players for Each Player Type") + 
  labs(x = "Player Type") + 
  scale_fill_discrete(name = "Active player")
```


Below is a histogram of the distribution of numerical variable wins. The plot shows that the probability of winning decreases as the number of wins increases. In other words, it is harder to win more games.
```{r}
#create histogram to present the distribution of numerical variable wins.
g <- ggplot(goalie_all, aes(x = wins, ..density..))
g + geom_histogram(bins = 150) +
  labs(x = "Wins", y = "Density") +
  ggtitle("Histogram for Wins") +
  geom_density(col = "red", lwd = 3, adjust = 0.4)
```

Below is a boxplot showing the number of wins for each conference. The Eastern conference has more wins.
```{r}
#create a new dataset conf_totals
conference <- stats(NULL)[ , c("id", "conference.name")]
totals <- total_list %>% 
  filter(gameTypeId == 2, teamId %in% c(1:26, 28:30, 52:55)) %>% 
  select(c(14, 26, 27, 30))
totals <- totals %>% rename(id = teamId)
conference <- conference %>% rename(Conference = conference.name)
conf_totals<- merge(totals, conference, by = "id")
#snapshot of the dataset conf_totals
head(conf_totals)

#create boxplot for number of wins for each conference
g <- ggplot(conf_totals, aes(x = Conference, y = wins)) 
g + geom_boxplot() + 
  geom_jitter(aes(color = Conference)) + 
  ggtitle("Boxplot for Wins") +
  labs(x = "Conference", y = "Wins")
```

Below is a boxplot showing the number of losses for each conference. Similar to the boxplot for wins, the Eastern conference has more losses.
```{r}
#create boxplot for number of losses for each conference
g <- ggplot(conf_totals, aes(x = Conference, y = losses)) 
g + geom_boxplot() + 
  geom_jitter(aes(color = Conference)) + 
  ggtitle("Boxplot for Losses") +
  labs(x = "Conference", y = "Losses")
```

The scatterplot below illustrates how wins and losses are related. We can see that more wins are accompanied by more losses. This is true for both conferences.
```{r}
#create scatter plot to show the relationship between wins and losses
g <- ggplot(conf_totals, aes(x = wins, y = losses, group = Conference)) 
g+ geom_point(aes(color = Conference), position = "jitter") + 
   geom_smooth(aes(group = Conference), method = lm, formula = y ~ x, col = "green") + 
   ggtitle("Wins vs Losses")
```

