---
title: "Project 1"
author: "Zichang Xiang"
date: "6/16/2021"
purpose: create vignette for reading and summarizing data from NHL API
output:  
  github_document:  
    toc: true
---
### Reading and summarizing data from NHL API
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Required Packages
jsonlite, RCurl, rmarkdown, httr, and ggplot2

```{r}
#load the required packages
library(rmarkdown)
library(dplyr)
library(jsonlite)
library(RCurl)
library(httr)
library(ggplot2)
```

### Functions to contact the NHL records API
#### Function to return franchise
```{r franchise}
#retrieve the info and convert to a list
get_franchise <- GET("https://records.nhl.com/site/api/franchise")
franchise_cont <- content(get_franchise, "text", encoding = "UTF-8")
franchise_json<- fromJSON(franchise_cont, flatten = TRUE)
franchise_list <-as_tibble(franchise_json$data)
#create function franchise
franchise <- function(name) {
  if (is.null(name)){
  return(franchise_list)
  }else if (is.character(name)){
  return(franchise_list %>% 
           filter(fullName == name | teamAbbrev == name | teamCommonName == name |
                    teamPlaceName == name))
  }else
  return(franchise_list %>% filter(mostRecentTeamId == name))
}
#try to use function franchise
franchise(NULL)
```


#### Function to return total stats for every franchise
```{r total}
#retrieve info and convert to a list
get_total <- GET("https://records.nhl.com/site/api/franchise-team-totals")
total_cont <- content(get_total, "text", encoding = "UTF-8")
total_json<- fromJSON(total_cont, flatten = TRUE)
total_list <-as_tibble(total_json$data)
#create function team_totals
team_totals <- function(name) {
  if (is.null(name)){
  return(total_list)
  }else if (is.character(name)){
  return(total_list %>% filter(teamName == name | triCode == name))
  }else
  return(total_list %>% filter(franchiseId == name) )
}

#try to use function team_totals
team_totals("MTL")
team_totals(NULL)
```

#### Function to return season records
```{r season}
#create function season
season <- function(name) {
  if (is.null(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-season-records")
    get_season <- GET(base_url)
    season_cont <- content(get_season, "text", encoding = "UTF-8")
    season_json <- fromJSON(season_cont, flatten = TRUE)
    season_list <- as_tibble(season_json$data)
  return(season_list)
  }else if (is.numeric(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", name)
    get_season <- GET(base_url)
    season_cont <- content(get_season, "text", encoding = "UTF-8")
    season_json <- fromJSON(season_cont, flatten = TRUE)
    season_list <- as_tibble(season_json$data)
  return(season_list)
  }else{
    get_season <- GET("https://records.nhl.com/site/api/franchise-season-records")
    season_cont <- content(get_season, "text", encoding = "UTF-8")
    season_json <- fromJSON(season_cont, flatten = TRUE)
    season_list <- as_tibble(season_json$data)
  return(season_list %>% filter(franchiseName == name))
  }
}

#try to use function season
season("Montréal Canadiens")
season(NULL)
```

#### Function to return goalie records
```{r goalie}
#create function goalie
goalie <- function(name) {
  if (is.null(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records")
    get_goalie <- GET(base_url)
    goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
    goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
    goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list)
  }else if (is.numeric(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", name)
    get_goalie <- GET(base_url)
    goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
    goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
    goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list %>% filter(franchiseId == name))
  }else{
    get_goalie <- GET("https://records.nhl.com/site/api/franchise-goalie-records")
    goalie_cont <- content(get_goalie, "text", encoding = "UTF-8")
    goalie_json <- fromJSON(goalie_cont, flatten = TRUE)
    goalie_list <- as_tibble(goalie_json$data)
  return(goalie_list %>% filter(franchiseName == name | lastName == name | 
                                    firstName == name))
  }
}

#try to use function goalie
goalie(1)
goalie(NULL)
```

#### Function to return skater records
```{r}
#create function skater
skater <- function(name) {
  if (is.null(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records")
    get_skater <- GET(base_url)
    skater_cont <- content(get_skater, "text", encoding = "UTF-8")
    skater_json <- fromJSON(skater_cont, flatten = TRUE)
    skater_list <- as_tibble(skater_json$data)
  return(skater_list)
  }else if (is.numeric(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", name)
    get_skater <- GET(base_url)
    skater_cont <- content(get_skater, "text", encoding = "UTF-8")
    skater_json <- fromJSON(skater_cont, flatten = TRUE)
    skater_list <- as_tibble(skater_json$data)
  return(skater_list %>% filter(franchiseId == name))
  }else{
    get_skater<- GET("https://records.nhl.com/site/api/franchise-skater-records")
    skater_cont <- content(get_skater, "text", encoding = "UTF-8")
    skater_json <- fromJSON(skater_cont, flatten = TRUE)
    skater_list <- as_tibble(skater_json$data)
  return(skater_list %>% filter(franchiseName == name | lastName==name | 
                                    firstName == name))
  }
}

#try to use function skater
skater(1)
skater(NULL)
```

#### Function to return detail records
```{r detail}
#create function detail
detail <- function(name) {
  if (is.null(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-detail?")
    get_detail <- GET(base_url)
    detail_cont <- content(get_detail, "text", encoding = "UTF-8")
    detail_json <- fromJSON(detail_cont, flatten = TRUE)
    detail_list <- as_tibble(detail_json$data)
  return(detail_list)  
  }else if (is.numeric(name)){
    base_url <- paste0("https://records.nhl.com/site/api/franchise-detail?cayenneExp=mostRecentTeamId=", name)
    get_detail <- GET(base_url)
    detail_cont <- content(get_detail, "text", encoding = "UTF-8")
    detail_json <- fromJSON(detail_cont, flatten = TRUE)
    detail_list <- as_tibble(detail_json$data)
  return(detail_list %>% filter(mostRecentTeamId == name))
  }else{
    get_detail <- GET("https://records.nhl.com/site/api/franchise-detail")
    detail_cont <- content(get_detail, "text", encoding = "UTF-8")
    detail_json <- fromJSON(detail_cont, flatten = TRUE)
    detail_list <- as_tibble(detail_json$data)
  return(detail_list %>% filter(teamAbbrev == name | teamFullName == name))
  }
}

#try to use function detail
detail("Vancouver Canucks")
detail(NULL)
```

### Function to contact the NHL stats API
```{r stats}
#create function stats
stats <- function(name) {
  if (is.null(name)){
    full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams",
                       "?expand=team.stats")
  }else{full_url <- paste0("https://statsapi.web.nhl.com/api/v1/teams/", name,
                           "/?expand=team.stats")}
    get_stats <- GET(full_url)
    stats_cont <- content(get_stats, "text", encoding = "UTF-8")
    stats_json <- fromJSON(stats_cont, flatten = TRUE)
    stats_list <- as_tibble(stats_json$teams)
  return(stats_list)
}

#try to use function stats
stats(NULL)
```

### Wrapper function to call other functions
```{r}
#create wrapper function to call other functions
wrapper <- function(fun, name){
  if (fun == "franchise"){
    return(franchise(name))
  }else if (fun == "team_totals"){
    return(team_totals(name))
  }else if (fun == "season"){
    return(season(name))
  }else if (fun == "goalie"){
    return(goalie(name))
  }else if (fun == "skater"){
    return(skater(name))
  }else if (fun == "detail"){
    return(detail(name))
  }else if (fun == "stats"){
    return(stats(name))
  }else{
    stop("Type the correct function!")
  }
}

#try to use function wrapper
wrapper("stats", NULL)
wrapper("franchise", NULL)
```

### Basic exploratory data analysis
#### Combine data from two endpoints
```{r}
#create new data frame goalie_all
goalie_all <- goalie(NULL)[ , c(1:8)]
goalie_all <- goalie_all %>% mutate(Type = "goalie")
#view the data frame
goalie_all

#create new data frame skater_all
skater_all <-skater(NULL)[, c(1:2, 4:8, 10)]
skater_all <- skater_all %>% mutate(Type = "skater")
#view the data frame
skater_all

#create new data frame goalie_skater_all
goalie_skater_all <- rbind(goalie_all, skater_all)
goalie_skater_all <- goalie_skater_all[order(goalie_skater_all$id), ]
#view the data frame
goalie_skater_all
```

#### Create at least two new variables  
Here we create three new variables loss_rate, win_rate, and tie_rate
```{r}
#create new data frame goalie_all
goalie_all <- goalie(NULL)[ , c(1:9, 28:29)]
#view the data frame
goalie_all

#create three new variables loss_rate, win_rate, and tie_rate
goalie_all <- goalie_all %>% mutate(loss_rate = goalie_all$losses/goalie_all$gamesPlayed, 
                                    win_rate = goalie_all$wins/goalie_all$gamesPlayed, 
                                    tie_rate = goalie_all$ties/goalie_all$gamesPlayed)
#view the data frame
goalie_all
```

#### Create contingency tables  
Table 1 below shows the relationship between the type of player and the number of active players. As we can see from the table, there are more non-active players than active players. This is true for both goalies and skaters.

In the second table, we can find the number of teams for each combination of division and time zone. We can see from the table that teams in division Honda West and in division Scotia North are from four different time zones, while teams in division MassMutual East fall into only one time zone EDT.
```{r}
#create table to summarize the relationship between player type and the number of active players
table(goalie_skater_all$Type, goalie_skater_all$activePlayer)

#create table to summarize the relationship between divisions and timeZones
table(stats(NULL)$division.name, stats(NULL)$venue.timeZone.tz)
```

#### Create numerical summaries  

```{r}
#quantiles and mean for wins in regular season
totals_regular <- team_totals(NULL) %>% filter(gameTypeId ==2) 
summary(totals_regular$wins)

#quantiles and mean for games played
goalie_skater_all %>% group_by(Type) %>% summarize(min = min(gamesPlayed), 
                                                   avg = round(mean(gamesPlayed),0), 
                                                   med = median(gamesPlayed), 
                                                   max = max(gamesPlayed))
```

#### Creates plots  
The bar plot below compares the number of players for each player type. As we can see, non-active skaters are the most common, while active goalies are the least common.
```{r}
#create bar plot to compare number of players for each player type
g <- ggplot(goalie_skater_all, aes(x = Type))
g + geom_bar(aes(fill = as.factor(activePlayer)), position = "dodge") + 
  ggtitle("Numer of Players for Each Player Type") + 
  labs(x = "Player Type") + 
  scale_fill_discrete(name = "Active player")
```


Below is a histogram of the distribution of numerical variable wins. The plot shows that the probability of winning decreases as the number of wins increases. In other words, it is harder to win more games.
```{r}
#create histogram to present the distribution of numerical variable wins.
g <- ggplot(goalie_all, aes(x = wins, ..density..))
g + geom_histogram(bins = 150) +
  labs(x = "Wins", y = "Density") +
  ggtitle("Histogram for Wins") +
  geom_density(col = "red", lwd = 3, adjust = 0.4)
```

Below is a boxplot showing the number of wins for each conference. The Eastern conference has more wins.
```{r}
#create a new dataset conf_totals
conference <- stats(NULL)[ , c("id", "conference.name")]
totals <- team_totals(NULL) %>% 
  filter(gameTypeId == 2, teamId %in% c(1:26, 28:30, 52:55)) %>% 
  select(c(14, 26, 27, 30))
totals <- totals %>% rename(id = teamId)
conference <- conference %>% rename(Conference = conference.name)
conf_totals<- merge(totals, conference, by = "id")
#view the dataset conf_totals
conf_totals

#create boxplot for number of wins for each conference
g <- ggplot(conf_totals, aes(x = Conference, y = wins)) 
g + geom_boxplot() + 
  geom_jitter(aes(color = Conference)) + 
  ggtitle("Boxplot for Wins") +
  labs(x = "Conference", y = "Wins")
```


The scatterplot below illustrates how wins and losses are related. We can see that more wins are accompanied by more losses. This is true for both conferences.
```{r}
#create scatter plot to show the relationship between wins and losses
g <- ggplot(conf_totals, aes(x = wins, y = losses, group = Conference)) 
g+ geom_point(aes(color = Conference), position = "jitter") + 
   geom_smooth(aes(group = Conference), method = lm, formula = y ~ x, col = "green") + 
   ggtitle("Wins vs Losses")
```

